A generalized definition of multidimensional item response theory parameters

### 2023-11-29 output/paper_draft/paper_M2PLM_parameters.Rmd
Rewrites "Test space and laten space" section content to match its new order in the draft

Thus, inner product $`r LS_BASIS_INNER_PROD`$ has $`r INNER_PROD_MAT_EQ`$ as
Gram matrix, i.e. $`r INNER_PROD_EQ`$, with $`r INNER_PROD_MAT_ELEMENT_EQ`$ as
the $`r INNER_PROD_MATRIX_INDEX`$-th element of $`r INNER_PROD_MATRIX`$. As
$`r TRANSFORM_MATRIX`$ (and consequently $`r TRANSFORM_MATRIX_TRANSP`$) is
invertible, so $`r INNER_PROD_MATRIX`$ is, and $`r INNER_PROD_MATRIX_INV_EQ`$.

@@

Before proceeding further, we need to make a clarification
regarding the representation of the items in the latent space.
Let's consider the model given by Equation \@ref(eq:M2PL-formulation),
where the latent traits and the parameters $`r DISCR_VECTOR`$ are expressed
in the same basis $`r LS_BASIS`$ in their respective spaces.
The coordinates of $`r DISCR_VECTOR`$ are then $`r DISCR_COORDS`$.
Since the response probability must be invariant with respect to
a change of basis, $`r DISCR_COORDS`$ must also be transformed
into $`r DISCR_STD_COORDS`$, such that $`r INVARIANCE_EQ`$;
therefore, $`r DISCR_STD_COORDS_EQ`$.
This shows that first a transformation through $`r INNER_PROD_MATRIX_INV`$
is performed and then a change to the standard basis,
so the $`r DISCR_STD_COORDS`$ are expressed in the standard basis.
The norm of $`r DISCR_VECTOR`$ must then fulfill

\begin{equation}
  `r DISCR_VECTOR_MODULE_EQ`.                    (\#eq:discr-param-squared-norm)
\end{equation}

Therefore, an inner product $`r TEST_SPACE_INNER_PROD`$ with Gram matrix
$`r INNER_PROD_MATRIX_INV`$ can be defined, such that $`r INNER_PROD_DISCR_EQ`$
keeps the norm of $`r DISCR_VECTOR`$ invariant.
The discrimination vectors are thus represented in their own space
$`r TEST_SPACE`$ with basis $`r TEST_SPACE_BASIS`$,
which we refer to as the *test space* [@zhang_theoretical_1999].
The test space is in general different from the latent space $`r LATENT_SPACE`$,
although there is a bijective relationship between the two sets of spaces.

@@

The slope in direction $`r DIR_ANGLE_VEC`$ is given by

@@

\begin{equation}
  \begin{split}
    `r DIR_COS_AS_DIR_COS_ORTH_EQ` \\
    `r TRAIT_POLAR_COEFF_COS_ORTH_EQ`.
  \end{split}                                 (\#eq:direction-cosine-vectors-eq)
\end{equation}

@@

We can then define

so

>>>>

Thus, inner product $`r LS_BASIS_INNER_PROD`$ has $`r INNER_PROD_MAT_EQ`$ as
Gram matrix, i.e., $`r INNER_PROD_EQ`$, with $`r INNER_PROD_MAT_ELEMENT_EQ`$ as
the $`r INNER_PROD_MATRIX_INDEX`$-th element of $`r INNER_PROD_MATRIX`$. As
$`r TRANSFORM_MATRIX`$ (and consequently $`r TRANSFORM_MATRIX_TRANSP`$) is
invertible, so $`r INNER_PROD_MATRIX`$ is, and $`r INNER_PROD_MATRIX_INV_EQ`$.

@@

Consider now the M2PL model formulation given by Equation
\@ref(eq:M2PL-formulation), on one hand, and Equation \@ref(eq:change-basis) on
the other.
Since the response probability must be invariant with respect to
a change of basis,
we must define the coordinates $`r DISCR_STD_COORDS`$ of the parameter vector
$`r DISCR_VECTOR`$ in an orthonormal basis $`r TEST_SPACE_STD_BASIS`$,
such that $`r INVARIANCE_EQ`$.
If we assume that $`r DISCR_VECTOR`$ exists in vector space $`r TEST_SPACE`$,
with coordinates $`r DISCR_COORDS`$ in basis $`r TEST_SPACE_BASIS`$,
then we must define $`r DISCR_COORDS`$ such that

We want @reckase_difficulty_1985's definition of the multidimensional parameters
in rectangular coordinates to be a particular case of our more general
definition, so we can conveniently assume that the inner product in
$`r TEST_SPACE_STD_BASIS`$ is the dot product,
as we did with $`r LS_ORTH_BASIS`$.
The norm of $`r DISCR_VECTOR`$ thus fulfills

\begin{equation}
  `r DISCR_VECTOR_MODULE_EQ`.
\end{equation}

Therefore, an inner product $`r TEST_SPACE_INNER_PROD`$ with Gram matrix
$`r INNER_PROD_MATRIX_INV`$ can be defined, such that $`r INNER_PROD_DISCR_EQ`$
keeps the norm of $`r DISCR_VECTOR`$ invariant.

This shows that discrimination vectors are indeed represented in their own
vector space $`r TEST_SPACE`$ with basis $`r TEST_SPACE_BASIS`$.
We shall refer to this new space as the *test space* [@zhang_theoretical_1999].
Although the inverse relationship between their respective inner product
Gram matrices induces a bijective relationship between the two sets of spaces,
it is worth highlighting that test space $`r TEST_SPACE`$
and latent space $`r LATENT_SPACE`$ are different entities. Only in the
orthonormal case the two spaces share a common inner product
(i.e., the dot product), and thus can be mistaken.

@@

The slope in direction $`r DIR_ANGLE_VEC_ORG`$ is given by

@@

\begin{equation}
  \begin{split}
    `r DIR_COS_AS_DIR_COS_ORTH_EQ`, \\
    `r TRAIT_POLAR_COEFF_COS_ORTH_EQ`.
  \end{split}                                 (\#eq:direction-cosine-vectors-eq)
\end{equation}

@@

As $`r DISCR_VECTOR`$ is defined in test space $`r TEST_SPACE`$ with basis
$`r TEST_SPACE_BASIS`$, its algebraic vector values correspond to its
coordinates, i.e., $`r DISCR_VECTOR_COORDS_EQ`$. Therefore, by Equation
\@ref(eq:discr-vector-std),
